A general Nullstellensatz for generalized spaces

### 2019-05-04 paper-qcoh.tex
typo

% XXX: mention internal language in general, and SDG?
\paragraph{Related work} The topos-theoretic Nullstellensatz is related to
several precursors. A corollary of the Nullstellensatz is that, over the
first-order theory validated by~$U_\TT$, any first-order formula is in fact
logically equivalent to a geometric formula. This corollary has already been
observed by Butz and Johnstone~\cite[Lemma~4.2]{butz-johnstone:first-order}. At
that point, a characterisation of the first-order formulas in the general case,
of the form as in Theorem~\ref{thm:characterization0}, was still missing.

@@

From the point of view of theory, the appropriate context for the
Nullstellensatz should be determined. We formulate the Nullstellensatz in the
context of geometric theories and their classifying Grothendieck toposes, but
there should be analogues for the Nullstellensatz in other contexts where it
makes sense to speak of ``classifying gadgets'', and these analogues should
shed light on the topos-theoretic version and ideally even explain it from deeper
principles. In particular, preliminary computations indicate that there is a
version of the Nullstellensatz for arithmetic universes, the predicative cousin
of toposes introduced by Joyal~\cite{XXX} and recently an important object of
consideration by Maietti and Vickers~\cite{XXX}.
\medskip

@@

To verify that~(3) implies~(4), we can mimic the usual proof of this fact in
the internal language of~$\E$: There is a variant of the syntactic
site of~$\TT$ which is built using only coherent sequents and finitary
derivability~\cite[Section~1.4]{caramello:tst}. The sheaf topos over this site
is another model for the classifying topos of~$\TT$, and still validates, like
any Grothendieck topos, full infinitary logic. Hence, if~$\sigma$
is~$\TT$-derivable by a proof tree of arbitrary shape, then~$\sigma$ holds in
this model of the classifying topos. By the analogue of
Proposition~\ref{prop:basic-truth} for this model, we obtain
that~$\sigma$ is~$\TT$-derivable by a proof tree of finite shape.

>>>>

% XXX: mention internal language in general, and SDG?
\paragraph{Related work} The topos-theoretic Nullstellensatz is related to
several precursors. A corollary of the Nullstellensatz is that, over the
first-order theory validated by~$U_\TT$, any first-order formula is in fact
logically equivalent to a geometric formula. This corollary has already been
observed by Butz and Johnstone~\cite[Lemma~4.2]{butz-johnstone:first-order}. At
that point, a characterization of the first-order formulas in the general case,
of the form as in Theorem~\ref{thm:characterization0}, was still missing.

@@

From the point of view of theory, the appropriate context for the
Nullstellensatz should be determined. We formulate the Nullstellensatz in the
context of geometric theories and their classifying Grothendieck toposes, but
there should be analogues of the Nullstellensatz in other contexts where it
makes sense to speak of ``classifying gadgets'', and these analogues should
shed light on the topos-theoretic version and ideally even explain it from deeper
principles. In particular, preliminary computations indicate that there is a
version of the Nullstellensatz for arithmetic universes, the predicative cousin
of toposes introduced by Joyal~\cite{XXX} and recently an important object of
consideration by Maietti and Vickers~\cite{XXX}.
\medskip

@@

To verify that~(3) implies~(4), we can mimic the usual proof of that for
coherent sequents, being derivable in geometric logic implies being derivable
in coherent logic, in
the internal language of~$\E$: There is a variant of the syntactic
site of~$\TT$ which is built using only coherent sequents and finitary
derivability~\cite[Section~1.4]{caramello:tst}. The sheaf topos over this site
is another model for the classifying topos of~$\TT$, and still validates, like
any Grothendieck topos, full infinitary logic. Hence, if~$\sigma$
is~$\TT$-derivable by a proof tree of arbitrary shape, then~$\sigma$ holds in
this model of the classifying topos. By the analogue of
Proposition~\ref{prop:basic-truth} for this model, we obtain
that~$\sigma$ is~$\TT$-derivable by a proof tree of finite shape.
